Using Cell Phone Keyboards is (\NP) Hard

### 2010-01-16 article.tex
Wording changes

Before we get any farther, let us sketch the basic problem that we will keep
revising and revisiting throughout this paper.
\begin{prob}[{\sc MinimumKeystrokes}]~\\
{\sc Instance}: A set of letters corresponding to an alphabet $A$ $(|A| =
n)$, a number of keys $k$, an input method $IN$, and a set of
tuples of words and frequencies $W$.  The frequencies in $W$ are integers,
and the words are made up of solely of elements of $A$.  We will
treat $IN$ as a function which, given a partition of $A$ and a word
$w$, returns how many keystrokes are required to type $w$.~\\
        ~\\
{\sc Question}: What is the best partition of $A$ into $k$ sets, such that the
total number of keystrokes to type every word in $W$ its associated frequency
times is minimized?  Equivalently, what is the partition of $P$ of $A$ ($|P| = k$) that
will minimize
$$\sum_{(w,f)\in W} f*IN(w,P)$$
\label{probtemplate}
\end{prob}

Over the course of this paper we consider three different real-world
schemes for $IN$ (basic typing, T9, and predictive T9), and for each
variant that is proven \NP-hard, we consider the restriction on $P$ that
requires that we keep the alphabet in alphabetical order.  In all cases
where it is computationally feasible we provide results for the case of
the English language, on 8-key keyboards, using the British National
Corpus\cite{bnc} (BNC).  One feature of note is that no matter what scheme we
use, the problem is trivial if the number of keys is not smaller than the
number of letters in the alphabet ($k \ge |A|$).  Using tiny keyboards is only
(computationally) difficult when the keyboards don't have enough keys.

@@

\begin{prob}[{\sc BasicCellPhoneTyping}]~\\
{\sc Instance}: A set of letters corresponding to an alphabet $A$ $(|A| =
n)$, a number of keys $k$, and a set of
tuples of words and frequencies $W$.  The frequencies in $W$ are integers,
and the words are made up of solely of elements of $A$. \\
        ~\\
{\sc Question}: What is the best partition $P$ of $A$ into $k$ sequences, such
that the total number of button presses to type every word in $W$ its
associated frequency times is minimized?  The number of button presses is equal
to the order of the letter in the sequence assigned to a given key.  For
example, in the key $2\to\{abc\}$, $a$ requires one keystroke and $c$ requires
3.  Equivalently, if we denote the number of button presses required to type
letter $a$ with partition $P$ as $IN_P(a)$, then we want to find the $P$ that
minimizes $$\sum_{(w,f) \in W}\sum_{c\in w} IN_P(c) * f$$
\label{bcpt}
\end{prob}

>>>>

Before we get any farther, let us sketch the basic problem that we will keep
revising and revisiting throughout this paper.
\begin{prob}[{\sc MinimumKeystrokes}]~\\
{\sc Instance}: A set of letters corresponding to an alphabet $A$ $(|A| =
n)$, a number of keys $k$, an input method $\mathrm{IN}$, and a set of
tuples of words and frequencies $W$.  The frequencies in $W$ are integers,
and the words are made up of solely of elements of $A$.  We will
treat $IN$ as a function which, given a partition of $A$ and a word
$w$, returns how many keystrokes are required to type $w$.~\\
        ~\\
{\sc Question}: What is the best partition of $A$ into $k$ sets, such that the
total number of keystrokes to type every word in $W$ its associated frequency
times is minimized?  Equivalently, what is the partition of $P$ of $A$ ($|P| = k$) that
will minimize
$$\sum_{(w,f)\in W} f*\mathrm{IN}(w,P)$$
\label{probtemplate}
\end{prob}

Over the course of this paper we consider three different real-world
schemes for $\mathrm{IN}$ (basic typing, T9, and predictive T9), and for each
variant that is proven \NP-hard, we consider the restriction on $P$ that
requires that we keep the alphabet in alphabetical order.  In all cases
where it is computationally feasible we provide results for the case of
the English language, on 8-key keyboards, using the British National
Corpus\cite{bnc} (BNC).  One feature of note is that no matter what scheme we
use, the problem is trivial if the number of keys is not smaller than the
number of letters in the alphabet ($k \ge |A|$).  Using tiny keyboards is only
(computationally) difficult when the keyboards don't have enough keys.

@@

\begin{prob}[{\sc BasicCellPhoneTyping}]~\\
{\sc Instance}: A set of letters corresponding to an alphabet $A$ $(|A| =
n)$, a number of keys $k$, and a set of
tuples of words and frequencies $W$.  The frequencies in $W$ are integers,
and the words are made up of solely of elements of $A$. \\
        ~\\
{\sc Question}: What is the best partition $P$ of $A$ into $k$ sequences, such
that the total number of button presses to type every word in $W$ its
associated frequency times is minimized?  The number of button presses is equal
to the order of the letter in the sequence assigned to a given key.  For
example, in the key $2\to\{abc\}$, $a$ requires one keystroke and $c$ requires
3.  Equivalently, if we denote the number of button presses required to type
letter $a$ with partition $P$ as $\mathrm{IN}_P(a)$, then we want to find the $P$ that
minimizes $$\sum_{(w,f) \in W}\sum_{c\in w} \mathrm{IN}_P(c) * f$$
\label{bcpt}
\end{prob}
